---
title: "team_git_star_assignment4"
author: "ZIYAO WANG (Billy)"
date: "2021/5/15"
output:
  bookdown::html_document2
keep_md: true  
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

#loading packages

library(tidyverse) 
library(plotly)
library(here)
library(readxl)
library(kableExtra)
```
Research questions:

Section 1 - Table 1
- [ ] For each LGA , which police region recorded maximum incidents?
- [ ] How did the trend of incidents recorded in those police regions varied across the years.

# Section 2

## For each LGA , which offence_subdivision, recorded maximum incidents
```{r read-data, include=FALSE}
criminaldata <- read_excel("data/Data_Tables_LGA_Criminal_Incidents_Year_Ending_December_2020.xlsx", sheet = 4)
```
```{r datacleaning, include=FALSE}
criminaluse <- criminaldata %>% 
  select(Year,
         `Local Government Area`,
         `Offence Subdivision`,
         `Incidents Recorded`)
```
```{r vis1, fig.width=9, fig.cap="The offence_subdivision of maximum incidents in each LGA"}
  criminalfinal<- criminaluse %>% 
  group_by(`Local Government Area`, `Offence Subdivision` ) %>% 
    summarise(incidents = sum(`Incidents Recorded`)) 

data<- criminalfinal %>% arrange(desc(incidents)) %>%
slice(1) %>% 
ggplot(aes(`Local Government Area`,
       incidents,
       fill = `Offence Subdivision` )) +
  geom_col()+
   ggtitle("The offence_subdivision of maximum incidents in each LGA")+
  theme(axis.text.x = element_blank()) 
 ggplotly(data)
```
The figure \@ref(fig:vis1)  shows, theft was the most common crime in each region, with Melbourne in particular recording some 90,900 burglaries in the past five years.
```{r vis2, fig.width=10, fig.cap="Each offence_subdivision's incidents in each year"}
  criminalfinal2<- criminaluse %>% 
  group_by(Year,`Offence Subdivision` ) %>% 
    summarise(incidents = sum(`Incidents Recorded`)) %>% 
ggplot(aes(Year,
       incidents,
       color = `Offence Subdivision` )) +
  geom_line()+
   ggtitle("Each offence_subdivision's incidents in each year")
 ggplotly(criminalfinal2)
```
The figure \@ref(fig:vis2)  shows

- Theft cases have far exceeded other types of crime in the past five years, especially reaching a peak (about 156,000) in 2016, but the number has declined in the past two years.

- Burlary/Break and Enter (about 38,000) and Assault and Related Offence (about 37,000) rank second and third, respectively.




## For each LGA,Which offence_subgroup was most recorded in the previous found offence_subdivision's?
```{r datacleaning2, include=FALSE}
criminaluse2 <- criminaldata %>% 
  select(Year,
         `Offence Subgroup`,
         `Offence Subdivision`,
         `Incidents Recorded`) %>% 
  filter(`Offence Subdivision` == "B40 Theft")
```
```{r vis3, fig.width=10, fig.cap="Incidents of each offence_subgroup in most recorded offence_subdivision"}
  criminalfinal3<- criminaluse2 %>% 
  group_by(Year, `Offence Subgroup`) %>% 
    summarise(incidents = sum(`Incidents Recorded`)) %>% 
ggplot(aes(Year,
       incidents,
       fill = `Offence Subgroup` )) +
  geom_col()+
   ggtitle("Incidents of each offence_subgroup in most recorded offence_subdivision")
 ggplotly(criminalfinal3)
```
The figure \@ref(fig:vis3)  shows

- Among thefts, steal from a motor vehicle has the largest number during the last five years. In 2016, about 63,000 steal from a motor vehicle occurred, but the number of thefts has decreased in the past two years.

- Other Theft is second(around40000), and other offence_subgroup make up only a small fraction.



##  What was the trend of each offence_subgroup found over the years?
```{r datacleaning3, include=FALSE}
criminaluse3 <- criminaldata %>% 
  select(Year,
         `Offence Subgroup`,
         `Incidents Recorded`)
```
```{r vis4, fig.width=10, fig.cap="Incidents of each offence_subgroup"}
  criminalfinal4<- criminaluse3 %>% 
  group_by(Year, `Offence Subgroup`) %>% 
    summarise(incidents = sum(`Incidents Recorded`)) %>% 
  arrange(desc(incidents)) %>% 
  head(100) %>% 
ggplot(aes(Year,
       incidents,
       color = `Offence Subgroup` )) +
  geom_line()+
   ggtitle("Incidents of each offence_subgroup")
 ggplotly(criminalfinal4)
```
The figure \@ref(fig:vis4)  shows

- Most have seen a decline in crime over the past year, and almost all have seen a small spike in 2016.

- Steal from a motor vehicle in the category of theft has been in the first place in the past five years, with about 50,000 cases occurring every year, far exceeding other offence_subgroups.

- Other Theft is in second place (about 39,000 per year) and Criminal Damage is third (about 35,000 per year). However, the gap between these two types of crimes and steal from a notor vehicle has gradually widened since 2013.

## Conclusion:

- In each LGA, theft occurs most frequently in offence_subdivision and the incidence of thefts is far greater than other types of offence_subdivision.

- In the theft offence_subdivision, the vast majority are steal from a motor vehicle and other Theft.

- Of all offence_subgroups, Steal from a motor vehicle in all offence_subgroups, Other Theft and Criminal Damage were the top three, but the vast majority of offence_subgroups were up in 2016, but have declined in the last two years.

Section 3

- [ ]  For each Suburb/Town, which offence_subdivision, recorded maximum incidents
- [ ] For each Suburb/Town,Which offence_subgroup was most recorded in the previous found offence_subdivision's?

