---
title: "team_git_star_assignment4"
author: "ZIYAO WANG (Billy)"
date: "2021/5/15"
output: html_document
---

Research questions:

Section 1 - Table 1

- [ ] For each LGA , which police region recorded maximum incidents?
- [ ] How did the trend of incidents recorded in those police regions varied across the years.

Section 2

- [ ]  For each LGA , which offence_subdivision, recorded maximum incidents
- [ ] For each LGA,Which offence_subgroup was most recorded in the previous found offence_subdivision's?
- [ ] What was the trend of each offence_subgroup found over the years?

Section 3

- [ ]  For each Suburb/Town, which offence_subdivision, recorded maximum incidents
- [ ] For each Suburb/Town,Which offence_subgroup was most recorded in the previous found offence_subdivision's?



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, fig.align = "center", fig.path = "Figures/")
```

```{r loadlibrary, message=FALSE}
library(tidyverse)
library(readxl)
library(haven)
library(ggplot2)
library(kableExtra)
library(ggResidpanel)
library(bookdown)
```

```{r readdata}
dat <- read_excel("data/Data_Tables_LGA_Criminal_Incidents_Year_Ending_December_2020.xlsx", sheet=2)

dat <- dat %>% 
  select(c(Year, `Police Region`, `Local Government Area`, `Incidents Recorded`)) 

dat_new <- dat %>% write.csv("data/LGA_Criminal_data2020.csv")

dat1 <- read.csv("data/LGA_Criminal_data2020.csv")
```

2011-2020 the total numeber of incidents recorded for each police region.
```{r table1, message=FALSE}
dat_tot <- dat1 %>% 
  filter(`Local.Government.Area` == "Total") %>% 
  group_by(`Police.Region`) %>% 
  summarise(Total_Incidents = sum(`Incidents.Recorded`)) %>% 
  arrange(-Total_Incidents)

table1 <- dat_tot %>% 
  kable(caption = "Number of Incidents Recorded in different Police Region", align = 'c') %>% 
  kable_styling(bootstrap_options = c("striped", "hover"))

```


```{r}
dat_var <- dat1 %>% 
  filter(`Local.Government.Area` == "Total") %>% 
  group_by(Year, `Police.Region`) %>% 
  summarise(Total_Incidents = sum(`Incidents.Recorded`))

ggplot(dat_var, aes(x= Year, y = Total_Incidents),
       color = `Police.Region`)+
  geom_line()+
  theme_bw()+
  facet_wrap(~ `Police.Region`)
```






