---
title: "team_git_star_assignment4"
author: "ZIYAO WANG (Billy)"
date: "2021/5/15"
output: html_document
---

Research questions:

Section 1 - Table 1
- [ ] For each LGA , which police region recorded maximum incidents?
- [ ] How did the trend of incidents recorded in those police regions varied across the years.

Section 2

- [ ]  For each LGA , which offence_subdivision, recorded maximum incidents
- [ ] For each LGA,Which offence_subgroup was most recorded in the previous found offence_subdivision's?
- [ ] What was the trend of each offence_subgroup found over the years?

Section 3

- [ ]  For each Suburb/Town, which offence_subdivision, recorded maximum incidents
- [ ] For each Suburb/Town,Which offence_subgroup was most recorded in the previous found offence_subdivision's?

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = FALSE, message = FALSE,warning = FALSE)
```

```{r read-lib,include=FALSE}
library(tidyverse)
library(dplyr)
library(naniar)
library(ggplot2)
library(plotly)
library(here)
library(readxl)
```

```{r read-file,include=FALSE}
data3 <- read_excel(here::here("data/Data_Tables_LGA_Criminal_Incidents_Year_Ending_December_2020.xlsx"),sheet = 4)
```

